\documentclass[../fem.tex]{subfiles}

\begin{document}
\section{Localization}%
\label{sec:localization}

While it is possible to find the elements of the matricies in
(\ref{eq:mat_rep}) on the global domain, this proves to be exedingly difficult,
and so we will use a more generalizable method. Since we are able to selece our
basis function $\varphi_j$ we can carefully construct a set of basis functions
such that it is possible to split the domain into discrete elements. Where each
element is a mutualy exclusive subset of the domain. We can then implement a
generalization to construct the system of equations for each element. Then
using an algorithm we combine each system of equations for each element to
construct the global matricies, which can then be solved for our approximation.

We discretize the domain $\Omega$ into $N$ triangles. This process of
constructing the triangular decomposition of the domain is detailed in the
numerical computation section \ref{sec:mesh_generation},  Mesh Generation. For
now we assume that we have constructed a mesh of $N$ triangles, and we denote
the each of these triangles $E_e$, where $e=1,\ldots,N$. We then also define
the triangle verticies to be labeled $E_e^{(i)}$ where $i=1,2,3$.

In 1D this step is fairly trivial as all elements have the exact same
dimesnions, but in our 2D implementation, each element can have variable shapes
and sizes. Inorder to resolve this we construct a master space.

\subsection{Master Space}%
\label{sub:master_space}

The master space is a domain that we construct inorder to simplify the
construction of our local approximations. For our situation with a triangular
element we construct our global space such that the verticies of a triangle lie
on the points $(1,0,0)$, $(0,1,0)$, and $(0,,01)$. Then for each element in our
mesh, we construct the transformation matrix $T$ to transform between master
space and element space. We define our master element to be $\hat{E}$. An image
of this master element can be seen in figure \ref{fig:master_element}.

\begin{Figure}
   \begin{center}
     \begin{tikzpicture}[scale=2]
        \draw[thick,->] (0,0,0) -- (1.3,0,0) node[right]{$\lambda_1$};
        \draw[thick,->] (0,0,0) -- (0,1.3,0) node[above]{$\lambda_2$};
        \draw[thick,->] (0,0,0) -- (0,0,1.3) node[above]{$\lambda_3$};
        \draw (1,0,0) node[above right]{$(1,0,0)$}-- (0,1,0)
        node[right]{$(0,1,0)$} -- (0,0,1) node[below right]{$(0,0,1)$}-- cycle;
      \end{tikzpicture}
   \end{center}
   \captionof{figure}{Master element $\hat{E}$, in barycentric coordinate
   system, that is used over the variable local
 elements.}\label{fig:master_element}
\end{Figure}

This master element is in what is called a Barycentric coordinate system.
Conversion from our trilinear coordinates to the Barycentric coordinates can be
done by using the following equations
\begin{align*}
  \lambda_1 &=
  \frac{\left(Y_2-Y_3\right)\left(x-X_3\right)+\left(X_3-X_2\right)\left(y-Y_3\right)}{\left(Y_2-Y_3\right)\left(X_1-X_3\right)+\left(X_3-X_2\right)\left(Y_1-Y_3\right)}\\
  \lambda_2 &=
  \frac{\left(Y_3-Y_1\right)\left(x-X_3\right)+\left(X_1-X_3\right)\left(y-Y_3\right)}{\left(Y_2-Y_3\right)\left(X_1-X_3\right)+\left(X_3-X_2\right)\left(Y_1-Y_3\right)}\\
  \lambda_3 &= 1-\lambda_1-\lambda_2.
\end{align*}
Where $X_i$ is the $x$ component of the $i$th vertex of the element, the same
applies for the $y$ component. $x$, and $y$ is the $x$ and $y$ component of the
point in the triangle that we want to convert to the Barycentric coordinates.

Converting from Barycentric coordinates to trilinear coordinates uses the
following equations
\begin{align*}
  x &= \lambda_1 X_1+\lambda_2 X_2 + \lambda_3 X_3\\
  y &= \lambda_1 Y_1 +\lambda_2 Y_2 + \lambda_3 Y_3
\end{align*}

\subsection{Local Basis}%
\label{sub:local_basis}


We want to construct three local basis functions, that are defined only for the
current element. We aim to construct three local basis functions functions
$\varphi_1^{(e)}$, $\varphi_2^{(e)}$, and $\varphi_3^{(e)}$.  Each shape
function should be linear, such that at its cooresponding vertex the local
basis function is $1$ and at the other verticies the function should evaluate
to $0$.

\otodo{Provide better reasoning for this neccesity.}

Conviniently this is exactly what the barycentric coordinates of the master
element provide us with, thus we define our local basis functions to be
\begin{align}\label{eq:local_basis}
  &\resizebox{\columnwidth}{!}{$
    \varphi_1^{(e)}=
  \frac{\left(Y^{(e)}_2-Y^{(e)}_3\right)\left(x-X^{(e)}_3\right)+\left(X^{(e)}_3-X^{(e)}_2\right)\left(y-Y^{(e)}_3\right)}{\left(Y^{(e)}_2-Y^{(e)}_3\right)\left(X^{(e)}_1-X^{(e)}_3\right)+\left(X^{(e)}_3-X^{(e)}_2\right)\left(Y^{(e)}_1-Y^{(e)}_3\right)}
  $}\\
  &\resizebox{\columnwidth}{!}{$
    \varphi_2^{(e)}=
    \frac{\left(Y^{(e)}_3-Y^{(e)}_1\right)\left(x-X^{(e)}_3\right)+\left(X^{(e)}_1-X^{(e)}_3\right)\left(y-Y^{(e)}_3\right)}{\left(Y^{(e)}_2-Y^{(e)}_3\right)\left(X^{(e)}_1-X^{(e)}_3\right)+\left(X^{(e)}_3-X^{(e)}_2\right)\left(Y^{(e)}_1-Y^{(e)}_3\right)}
  $}\\
  &\varphi_3^{(e)}=1-\varphi_1^{(e)}-\varphi_2^{(e)}
\end{align}

Thus for every element, there are three local basis functions. A plot of these
local basis functions is shown in figure \ref{fig:local_basis}.

\begin{Figure}
   \begin{center}
     \begin{tikzpicture}[scale=3]
       \draw[thick] (0,0,0)node[below]{$E^{(e)}_1$} --
         (1,0,0)node[below]{$E^{(e)}_2$}
         --node[below]{$E_e$} (0.6, 0, 0.5)node[below]{$E^{(e)}_3$} -- cycle;
       \draw (0,1,0)node[above]{$\varphi_1^{(e)}$}  -- (1,0,0) -- (0.6, 0, 0.5)-- cycle;
       \draw (0,0,0) -- (1,1,0)node[above]{$\varphi_2^{(e)}$}  -- (0.6, 0, 0.5)-- cycle;
       \draw (0,0,0) -- (1,0,0) -- (0.6, 1, 0.5)node[above]{$\varphi_3^{(e)}$} -- cycle;
       \draw[dotted] (0,0,0) --node[left]{$1$} (0,1,0);
       \draw[dotted] (1,0,0) --node[right]{$1$} (1,1,0);
       \draw[dotted] (0.6,0,0.5) --node[right]{$1$} (0.6,1,0.5);
      \end{tikzpicture}
   \end{center}
   \captionof{figure}{Plot of local basis function $\varphi_1^{(e)}$,
   $\varphi_2^{(e)}$, and $\varphi_3^{(e)}$, on some arbatrairy element
 $E_e$.}\label{fig:local_basis}
\end{Figure}

\subsection{Global Basis}%
\label{sub:global_basis}

We now construct the global basis functions. We chose the global basis
functions to be associated with a vertex of the mesh, thus for $N$ mesh
verticies are $N$ global basis functions. We want the global basis functions be
be $1$ at their associated vertex, and at all other mesh verticies the global
basis function should be $0$. Inorder to achive this we construct the global
basis functions as a piecewise combination of $n$ different local basis
functions, where $n$ is the number of elements that share a givin vertex. There
is not a simple mathematical expression for the global basis function, as it
greatly depends on the triangular mesh, and the current vertex index. However,
an example of a global basis function is shown in figure
\ref{fig:global_basis}.

\begin{Figure}
   \begin{center}
      \begin{tikzpicture}
        \draw (0,2,0) -- (1,0,-2) -- (2,0,0) -- cycle;
        \draw (0,2,0) -- (2,0,0) -- (1,0,1) -- cycle;
        \draw (0,2,0) -- (1,0,1) -- (-2,0,1) -- cycle;
        \draw (0,2,0) -- (-2,0,1) -- (-1,0,-2) -- cycle;
        \draw (0,2,0) -- (-1,0,-2) -- (1,0,-2) -- cycle;

        \draw[thick] (0,0,0) -- (1,0,-2) -- (2,0,0) -- cycle;
        \draw[thick] (0,0,0) -- (2,0,0) -- (1,0,1) -- cycle;
        \draw[thick] (0,0,0) -- (1,0,1) -- (-2,0,1) -- cycle;
        \draw[thick] (0,0,0) -- (-2,0,1) -- (-1,0,-2) -- cycle;
        \draw[thick] (0,0,0) -- (-1,0,-2) -- (1,0,-2) -- cycle;

        \draw[dotted, thick] (0,0,0)node[below left]{$n$} -- (0,2,0) node[above]{$\varphi_n$};
      \end{tikzpicture}
   \end{center}
   \captionof{figure}{Gloabal basis function $\varphi_n$, demonstrating a
   potential shape of the global basis functions on a limited portion of a
 triangular mesh.}\label{fig:global_basis}
\end{Figure}

\subsection{Local System}%
\label{sub:local_system}

We use our definition of the local basis functions to apply in the construction
of an element specific system of equations from (\ref{eq:mat_rep}). Thus we
construct our local systems of equations in matrix form as
\begin{align*}
  \left(C^{(e)}+K^{(e)}\right)T^{(e)}=F^{(e)}.
\end{align*}

Here the elements of each matrix is only determined by the local basis function
of the element $e$, and the size of the matrix is equivalent to the number of
local degrees fo freedom, whihc in this case is two (One for each vertex of the
element). The elements of these matrices can be found through the following
equations
\begin{align}\label{eq:local_int}
  c_{ij}^{(e)}&=\rho
  c_p\int_{E_e}\varphi_i^{(e)}\vec{u}\cdot\nabla\varphi_j^{(e)}d\vec{x}\quad&i,j&=1,2,3\\
  k_{ij}^{(e)}&=\int_{E_e}\varphi_i^{(e)}\cdot\nabla\varphi_j^{(e)}d\vec{x}\quad&i,j&=1,2,3\\
  f_i^{(e)}&=\int_{E_e}\varphi_i^{(e)}fd\vec{x}\quad&i&=1,2,3.
\end{align}
Thus we only need to compute these local matrix elements, then construct the
global system utilizing all of the local matricies.

\subsection{Numerical Integration}%
\label{sub:numerical_integration}

As it can be seen in equation (\ref{eq:local_int}) we need to be able to
integrate over the element $E_e$. Inorder to achive this integration over a
triangular element, we implement the Barycentric coordinate systems, and a
method of numerical integration called Gaussian Quadrature.

By using the Barycentric coordinates, our integrals become significantly easier
to evaluate.
\begin{align*}
  I&=\int_{E_e}f(x,y)da\\
   &=\int_0^1\int_0^1f\left(\bar{x}, \bar{y}\right)d\lambda_1 d\lambda_2,
\end{align*}
where $\bar{x}$ and $\bar{y}$ are the $x$ and $y$ coordinates that are achived
from the coordinate transfer from Barycentric to trilinear.

Using this formulation of the integral, any numberical integration methods
should suffice. We will implement 2D Gaussian Quadrature, defined on the
triangle.

Gaussian Quadrature is a method that approximates the integral such that
\begin{align*}
  \int f(x)dx\approx \sum_{i=1}^nw_if(x_i),
\end{align*}
where $w_i$, and $x_i$ are specified weights and positions respectivly. For the
2D alternative, we can express this in the form
\begin{align*}
  \int_\gamma f(x,y) dxdy\approx \frac{1}{2}\sum_{i=1}^nw_if(x_i,y_i).
\end{align*}
Where $\gamma$ is the triangle with verticies $(0,0)$, $(1,0)$, and $(0,1)$.
This is the exact verticies that are provided by the triangle over $\lambda_1$
and $\lambda_2$. Thus we use the gaussian quadrature implementation in the
Barycentric coordinate system, and convert the coordinates into trilinear
coordinates inorder to evaluate whatever function at the point.

The methods for constructing the weights and positions of sample points for
this form of integration is out of the scope of this paper, and so will be
ignored. We utilize the weights and points constructed by \cite{GQ}.

\end{document}
