\documentclass[../fem.tex]{subfile}

\begin{document}
\section{Mesh Generation}%
\label{sec:mesh_generation2}

The mathematical theory of the mesh generation was discussed in section
\ref{sec:mesh_generation}. This section will describe the specifics of the
implementation, and the algorithms that are utilized for the construction of
the Delaunay triangulation of an arbitrary domain.

\subsection{Edge Flipping}%
\label{sub:edge_flipping}

We utilize an implementation of edge flipping algorithms developed by Sloan.
Sloan's edge flipping algorithm has implementation specifics for both
constrained and unconstrained Delaunay triangulation.

We will provide a short explanation of the method that is used for the
construction of our triangular mesh, but more detail can be found in
\cite{S_DT}\cite{S_CDT}. First Sloan's algorithm constructions the
unconstrained Delaunay triangulation then enforces the required edges into the
triangulation. So, we begin with the process for construction the unconstrained
triangulations.

\subsubsection{Construction of Delaunay Triangulation}%
\label{ssub:construction_of_delaunay_triangulation}

Using this method also provides the advantage of generating a triangle
adjacency list for each of the triangles. This allows for optimizations later
in the process of finite element analysis.

There are seven main stages of the construction of the Delaunay triangulation.
We also define $G$ as the set of points to be triangulated, and $N$ as the
number of points in $G$.

\begin{enumerate}[label=\arabic*.]
  \item Normalize coordinates of points.
  \item Sort points into bins.
  \item Establish the super triangle.
  \item Loop over each point, repeating 5-7
  \item Insert new point in triangulation.
  \item Initialize stack.
  \item Restore Delaunay triangulation.
\end{enumerate}

In stage 1, we scale all the points that will be used to construct the
triangulation between $0$ and $1$, this should be done such that all relative
positions of points are unchanged. This is done using
\begin{align*}
  \hat{x}=\frac{x-x_{min}}{d_{max}},\quad\hat{y}=\frac{y-y_{min}}{d_{max}},
\end{align*}
where
\begin{align*}
  d_{max}=\max\left\{x_{max}-x_{min},\ y_{max}-y_{min}\right\}.
\end{align*}
This will scale all coordinates between $0$ and $1$ but will maintain the
relative positions to all other points.

In stage 2, the normalized points are sorted into a set of bins. Each bin is
associated with a rectangular portion of global space that the points are
included in. We construction the bins such that each bin contains roughly
$\sqrt{N}$ points. The bins are then ordered such that adjacent bins are
numbered consecutively. This is done to improve the efficiency of later
searching of the triangular mesh.

In step 3, we construction a ``super triangle''. This triangle is a triangle of
three additional points, such that all points in $G$ are enclosed within the
super triangle.

In stage 4, we repeat stages 5-7 for every $P$ in $G$.

In step 5, we determine the triangle which encloses $P$(Since the super
triangle encloses all points, then this triangle must exist for $P$). We now
delete this triangle, and construct three new triangles, by connection $P$ to
the three vertices of the deleted triangle.

In step 6, we add up to three triangles that are adjacent to the edges opposite
$P$ to a stack(A \textit{stack} is a last in first out data structure, like a
stack of trays).

In step 7, while the stack is not empty we proceed as following

\begin{enumerate}[label=7.\arabic*.]
  \item Remove triangle from top of the stack.
  \item If $P$ is within the circumcircle of this triangle, then the adjacent
    triangle containing $P$ and the unstacked triangle form a convex
    quadrilateral whose diagonal is drawn in the non-optimal direction. Swap
    this diagonal edge.
  \item Add any new triangles which are opposite $P$ to the stack.
\end{enumerate}

After the stack is empty, then Delaunay triangulation has been restored.

Once every point has been added to the triangulation, then the completed
Delaunay triangulation can be returned.

\subsubsection{Construction of Constrained Delaunay Triangulation}%
\label{ssub:construction_of_constrained_delaunay_triangulation}

Now that a Delaunay triangulation has been constructed for our set of $G$
points, we now modify the triangulation so to ensure that certain edges are
present.

This process proceeds as follows

\begin{enumerate}[label=\arabic*.]
  \item Loop over each constrained edge.
  \item Find intersecting edges.
  \item Remove intersecting edges.
  \item Restore Delaunay triangulation.
  \item Remove superfluous triangles.
\end{enumerate}

In step 1, we loop over every edge to be constrained, and we define the edge by
the endpoints of the edge $V_i$ and $V_j$. Using this edge repeat steps 2-4.

In stage 2, we find all edges in the triangulation that intersect $V_i-V_j$.
Store all of these edges in a list. If our constrained edge is already present
in the triangulation, the go to step 1.

In step 3, repeat while there are still edges that intersect $V_i-V_j$.
\begin{enumerate}[label=3.\arabic*.]
  \item Remove an edge from the list of intersecting edges, call this edge
    $V_k-V_l$.
  \item If the triangles that share the edge $V_k-V_l$ do not form a strictly
    convex quadrilateral, then place the edge back on the list of edges and go
    to step 3. Else, swap this diagonal, and define the new diagonal as
    $V_m-V_n$. If $V_m-V_n$ still intersections $V_i-V_j$, then place it back
    on the list of new edges, otherwise place $V_m-V_n$ on a list of new edges.
\end{enumerate}

In step 4, repeat until no changes occur.
\begin{enumerate}[label=4.\arabic*.]
  \item Loop over each newly created edge, define each edge by $V_k-V_l$.
  \item If $V_k-V_l$ is equal to $V_i-V_j$, then skip to step $4.1$.
  \item If the two triangles that share the edge $V_k-V_l$ do not satisfy the
    Delaunay criterion, then swap the diagonal, and place the new diagonal on
    the list of new edges.
\end{enumerate}

In step 5, we need to remove all unwanted triangles. For this stage, we differ
from the process of Sloan. This is because for our implementation we desire the
ability to have holes in the mesh, and the simplistic method of removing
exterior triangles will not achieve this.

We use what is known as a triangle infection algorithm. Where we define a few
points that will infect the triangles which the points are within, then the
infection spreads to all adjacent triangles that are not separated by a
constrained edge. This process is repeated until no new triangles become
infected. Then all infected triangles are removed.

With this process, we need to define a point that is within every one of the
holes in the mesh, as the initial infection point, then the three vertices of
the super triangle are also used as initial infection points.

After all infected triangles are removed, we have constructed our constrained
Delaunay triangulation, which can then be used for FEA.

\subsection{Mesh Refinement}%
\label{sub:mesh_refinement2}

The raw result of our constrained Delaunay triangulation algorithm may be
sufficient in a few situations, but for most purposes, it returns sub-optimal
meshes. Because of this, we must implement a mesh refinement algorithm to
improve upon the quality of our mesh.

There are two main mesh refinement algorithms, Ruppert's algorithm
\cite{R_REF}, and Chew's second algorithm \cite{C_REF}. These two algorithms
work with a similar principle, but we will focus on Chew's second algorithm. We
do this because of the advantages that Chew's second algorithm provides over
Ruppert's.

Mesh refinement is required because triangular elements along constrained
edges are currently forced to have the constrained edge and one of their edges.
This can cause these triangles to be skinny sliver triangles, ones that are
undesirable in our final mesh.

Using the definitions from section \ref{sub:mesh_refinement}we are able to
implement the mesh refinement algorithm.  Chew's second algorithm proceeds as
follows.

\begin{enumerate}[label=\arabic*.]
  \item Grade any triangles that are currently ungraded. A triangle only passes
    if it is \textit{nice}(Def \ref{def:nice_tri}).
  \item If all triangles pass then Halt. Otherwise select the larges triangle
    that fails $\Delta$, and determine its circumcenter $c$.
  \item Traverse the triangulation from any vertex of $\Delta$ in the direction
    of $c$ until either running into a source-edge or finding the triangle
    containing $c$.
  \item If the triangle containing $c$ was found then insert $c$ into the
    triangulation, and update the triangulation to be Delaunay. This process is
    similar to that of sec \ref{sub:edge_flipping}. Then go to step 2.
  \item If a source edge was encountered, then split the source edge into two
    equal sized edges and update the triangulation. Let $l$ be the length of
    the new edges. Delete each circumcenter-vertex(vertices that are not part
    of the original triangulation) that is within $l$(line-of-sight distance,
    where a source-edge means that a vertex is infinitely far away) of the new
    vertex. Then go to step 2.
\end{enumerate}

Using this algorithm it is possible to construct triangulations with a required
minimum angle and force a size function. This provides the ability of
guaranteed \textit{nice} triangular meshes for any desired input domain.

Both Chew's second algorithm and Ruppert's algorithm have proven termination
for minimum angles below a certain point. For Rupert's algorithm, any minimum
angle below $20.7^{\circ}$ is guaranteed to halt. While Chew's second algorithm
is guaranteed to halt for angles below $28.6^{\circ}$ and will often succeed
with angles below $34^{\circ}$. Because of this improvement on the guaranteed
minimum angle is why our focus is on Chew's algorithm for mesh refinement.

% \begin{Figure}
%    \begin{center}
%      \begin{minipage}{0.4\textwidth}
%        \begin{center}
%          % \input{figures/delaunay_1.tikz}
%          (a)
%        \end{center}
%      \end{minipage}
%      \begin{minipage}{0.4\textwidth}
%        \begin{center}
%          % \input{figures/delaunay_4.tikz}
%          (b)
%        \end{center}
%      \end{minipage}
%    \end{center}
%    \captionof{figure}{These two images demonstrate the progressive refinement
%    of a mesh depending of the refinement restrictions of angle and area. (a)
%  $60$ triangles, (b) $132$ triangles.}
% \end{Figure}


\subsection{Implementation}%
\label{sub:implementation}

There are many possible implementations of this element discretization process.
Several of the original papers provide FORTRAN code, that can be used in
implementations. We transcribed Sloan's FORTRAN code for his implementation of
Constrained Delaunay Triangulation construction into C++. However, for mesh
refinement, we utilized the open source software Triangle\cite{T_SOF}, which
implements the Delaunay triangulation construction and mesh refinement.
Triangle implements several algorithms, we chose to utilize the divide and
conquer algorithm, developed by Chew.

\end{document}
